---
title: "Thoracic Surgery Data Week 10"
author: "Gillian Tatreau"
date: "2022-11-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r SET UP, include=FALSE}
library(foreign)
library(caTools)

setwd("/Users/gillian/Documents/Bellevue Grad Program/Fall 2022/DSC520/DSC520 Repo")

file_name <- "/Users/gillian/Documents/Bellevue Grad Program/Fall 2022/DSC520/DSC520 Repo/ThoraricSurgery.arff"
surgery <- read.arff(file_name) 
head(surgery)

split <- sample.split(surgery, SplitRatio = 0.8)
split
train <- subset(surgery, split == "TRUE")
test <- subset(surgery, split == "FALSE")

colSums(is.na(surgery))

model1 <- glm(Risk1Yr ~ AGE, data = train, family = binomial())
model2 <- glm(Risk1Yr ~ AGE + DGN, data = train, family = binomial())
model3 <- glm(Risk1Yr ~ AGE + DGN + PRE4, data = train, family = binomial())
model4 <- glm(Risk1Yr ~ AGE + DGN + PRE4 + PRE5, data = train, family = binomial())
model5 <- glm(Risk1Yr ~ AGE + DGN + PRE4 + PRE5 + PRE6, data = train, family = binomial())
model6 <- glm(Risk1Yr ~ AGE + DGN + PRE4 + PRE5 + PRE6 + PRE7, data = train, family = binomial())
model7 <- glm(Risk1Yr ~ AGE + DGN + PRE4 + PRE5 + PRE6 + PRE7 + PRE8, data = train, family = binomial())
model8 <- glm(Risk1Yr ~ AGE + DGN + PRE4 + PRE5 + PRE6 + PRE7 + PRE8 + PRE9, data = train, family = binomial())
model9 <- glm(Risk1Yr ~ AGE + DGN + PRE4 + PRE5 + PRE6 + PRE7 + PRE8 + PRE9 + PRE10, data = train, family = binomial())
model10 <- glm(Risk1Yr ~ AGE + DGN + PRE4 + PRE5 + PRE6 + PRE7 + PRE8 + PRE9 + PRE10 + PRE11, data = train, family = binomial())
model11 <- glm(Risk1Yr ~ AGE + DGN + PRE4 + PRE5 + PRE6 + PRE7 + PRE8 + PRE9 + PRE10 + PRE11 + PRE14, data = train, family = binomial())
model12 <- glm(Risk1Yr ~ AGE + DGN + PRE4 + PRE5 + PRE6 + PRE7 + PRE8 + PRE9 + PRE10 + PRE11 + PRE14 + PRE17, data = train, family = binomial())
model13 <- glm(Risk1Yr ~ AGE + DGN + PRE4 + PRE5 + PRE6 + PRE7 + PRE8 + PRE9 + PRE10 + PRE11 + PRE14 + PRE17 + PRE19, data = train, family = binomial())
model14 <- glm(Risk1Yr ~ AGE + DGN + PRE4 + PRE5 + PRE6 + PRE7 + PRE8 + PRE9 + PRE10 + PRE11 + PRE14 + PRE17 + PRE19 + PRE25, data = train, family = binomial())
model15 <- glm(Risk1Yr ~ AGE + DGN + PRE4 + PRE5 + PRE6 + PRE7 + PRE8 + PRE9 + PRE10 + PRE11 + PRE14 + PRE17 + PRE19 + PRE25 + PRE30, data = train, family = binomial())
model16 <- glm(Risk1Yr ~ AGE + DGN + PRE4 + PRE5 + PRE6 + PRE7 + PRE8 + PRE9 + PRE10 + PRE11 + PRE14 + PRE17 + PRE19 + PRE25 + PRE30 + PRE32, data = train, family = binomial())

summary(model1)
summary(model2)
summary(model3)
summary(model4)
summary(model5)
summary(model6)
summary(model7)
summary(model8)
summary(model9)
summary(model10)
summary(model11)
summary(model12)
summary(model13)
summary(model14)
summary(model15)
summary(model16)

model17 <- glm(Risk1Yr ~ PRE14 + PRE9 + PRE17 + PRE30 + PRE4 + PRE5 + PRE6 + AGE, data = train, family = binomial())

# compare model 1 and model 15
modelChi1 <- model1$deviance - model15$deviance
chidf1 <- model1$df.residual - model15$df.residual
chisq.prob1 <- 1 - pchisq(modelChi1, chidf1)
modelChi1; chidf1; chisq.prob1

# compare model 1 and model 17
modelChi2 <- model1$deviance - model17$deviance
chidf2 <- model1$df.residual - model17$df.residual
chisq.prob2 <- 1 - pchisq(modelChi2, chidf2)
modelChi2; chidf2; chisq.prob2

# compare model 17 and model 15
modelChi3 <- model17$deviance - model15$deviance
chidf3 <- model17$df.residual - model15$df.residual
chisq.prob3 <- 1 - pchisq(modelChi3, chidf3)
modelChi3; chidf3; chisq.prob3
```


# i. 
```{r echo=FALSE}
model15 <- glm(Risk1Yr ~ AGE + DGN + PRE4 + PRE5 + PRE6 + PRE7 + PRE8 + PRE9 + PRE10 + PRE11 + PRE14 + PRE17 + PRE19 + PRE25 + PRE30, data = train, family = binomial())
summary(model15)
```

# ii. 
According to the summary, the variables that had the greatest affect on survival rate (those with a p-value of less than 0.25) were PRE14, PRE9, PRE17, PRE30, PRE4, PRE5, PRE6, and AGE. Therefore, the model that would be the most accurate would include just those variables in the order from most significant to least significant p-values. 
```{r echo=FALSE}
model17 <- glm(Risk1Yr ~ PRE14 + PRE9 + PRE17 + PRE30 + PRE4 + PRE5 + PRE6 + AGE, data = train, family = binomial())
summary(model17)
```

# iii. 
The model was able to predict the correct value for the training data with approximately an 85% accuracy. The model was able to predict the correct value for the testing data with approximately an 82% accuracy.
```{r echo=TRUE}
# accuracy for training data
res <- predict(model17, train, type = "response")
res

confmatrix <- table(Actual_Value = train$Risk1Yr, Predicted_Value = res > 0.5)
confmatrix

(confmatrix[[1,1]] + confmatrix[[2,2]]) / sum(confmatrix)

```

```{r echo=TRUE}
# accuracy for testing data
res <- predict(model17, test, type = "response")
res
confmatrix <- table(Actual_Value = test$Risk1Yr, Predicted_Value = res > 0.5)
confmatrix

(confmatrix[[1,1]] + confmatrix[[2,2]]) / sum(confmatrix)

```


## Code Appendix
```{r ref.label = knitr::all_labels(), echo = TRUE, eval = FALSE}
```



